---
title: "Assignment 1 Report"
author: "Harnaaz Sandhu"
date: "2025-01-07"
output: html_document
---
GitHub repository for this assignment- 
https://github.com/hhsandhu/BIOL432_Assignment1 

# Introduction

This report summarizes the results of generating limb measurements and estimating limb volumes for hypothetical organisms.

# Part 1: Generate Limb Measurements

I used the script `dataGenerato.R` to create a dataset of limb measurements. This script randomly generated species names, limb widths, limb lengths, and observer names, and saved them in a file called `measurements.csv`.

# Part 2: Estimate Limb Volumes

I used the script volumeEstimato.R to calculate limb volumes based on the limb width and length, and updated the measurements.csv file with a new column for Volume. 

Source the script to generate data
```{r, echo=TRUE}
source("dataGenerato.R")
```

Source the script to estimate limb volumes
```{r, echo=TRUE}
source("volumeEstimato.R")
```

# Part 3: Analyze the Data

```{r, echo=TRUE}
library(dplyr)
data <- read.csv("measurements.csv")
sorted_data <- data %>%
  arrange(Organism, Observer, Volume)
head(sorted_data)
```

Table showing the average limb volume for each species.
```{r, echo=TRUE}
average_volume <- data %>%
  group_by(Organism) %>%
  summarize(Average_Volume = mean(Volume))
average_volume
```

Table showing the number of observations for each combination of species and observer.
```{r, echo=TRUE}
observations <- data %>%
  group_by(Organism, Observer) %>%
  summarize(Count = n())
observations
```

Boxplot to compare the distribution of limb volumes for each species.
```{r, echo=TRUE}
library(ggplot2)

ggplot(data, aes(x = Organism, y = Volume)) +
  geom_boxplot() +
  theme_minimal() +
  labs(title = "Limb Volume by Species", x = "Species", y = "Volume")
```

Multi-panel histogram showing the distribution of limb volumes for each species
```{r, echo=TRUE}
library(ggplot2)

ggplot(data, aes(x = Volume)) +
  geom_histogram(binwidth = 10, fill = "skyblue", color = "black") +
  facet_wrap(~ Organism, scales = "free") +
  theme_minimal() +
  labs(
    title = "Distribution of Limb Volumes by Species",
    x = "Volume",
    y = "Frequency"
  )
```


